---
title: "Multiplots"
author: "Devin Pastoor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r}
library(PKPDmisc) # for binning and other functions
library(PKPDdatasets) # example dataset
library(ggplot2) #plotting
library(purrr) # dataset to handle lists
suppressMessages(suppressWarnings(library(dplyr)))
```

All packages (except for ggplot) are only available on github

```
devtools::install_github("dpastoor/PKPDdatasets",
              "dpastoor/PKPDmisc", 
              "hadley/purrr")

```

```{r}
dat <- sd_oral_richpk
```

Messy plot can't see individuals well

```{r}
ggplot(dat, aes(x = Time, y = Conc, group= ID)) + 
  geom_line() +
  facet_wrap(~ID, scales= "free")
```

What we'd like to be able to do is split up into 'bins' of specified numbers of individuals to then plot out multiple plots

```{r}
# this will automatically create a column of bins such that the
# specified number of individuals is in each bin, in this case
# 9 ids per bin
dat$IDBINS <- ids_per_plot(dat$ID)

# split the original dataset into subdatasets corresponding to each bin (list of dataframes)
split_dat <- dat %>% split(.[["IDBINS"]])
```

To handle plotting each subdataframe, you need to wrap your normal ggplot into a function. You will then apply this function with `map` to each subdataframe. So in this case it is just like a normal plot, but wrapped up in a function, which takes 1 argument (the dataframe) and outputs the plot

```{r}
p_conc_time <- function(df) {
ggplot(df, aes(x = Time, y = Conc, group= ID)) + 
  geom_line() +
  facet_wrap(~ID, scales= "free")
}
```


To apply the above function we use `map` from `purrr`

```{r}
split_dat %>% map(p_conc_time) 
```

notice no quotes needed on the function

