---
title: "NONMEM readers"
author: "Devin Pastoor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

NONMEM's formatting of output (appending tables), especially in the context of using `NSUB` has been cause for much frustration for parsing data. Not only does it break the delimiting, but it also has inconsistent formatting:

* no leading spaces for `TABLE` lines
* one leading space for column names
* two leading spaces for numerical values

While a number of fast-reading packages are available in R, including `data.table::fread()`, and the new `readr::read_csv`, these solutions cannot handle NONMEM's output gracefully, and therefore pure R solutions such as that of `xpose4::read.nm.tables()` must be used. 

This vignette introduces `PKPDmisc::read_nonmem()`, a c++ implementation to extend `readr` to handle nonmem tabular output by properly parsing rows, including:

* removal of `TABLE` lines
* proper trimming of whitespace

As shown below, this package is nearly 13 times faster than xpose's current implementation.

```{r}
library(PKPDmisc)
library(microbenchmark)
library(ggplot2)
```

To test, 2 larger datasets (~20 MB and ~120 MB respectively) are benchmarked against
current implementations.
```{r}
med_20mb_file <- "~/Repos/misc-datasets/nonmem_sim.vpc"
large_120mb_file <- "~/Repos/misc-datasets/nonmem_sim_big.vpc"
```

```{r}
med_rtnm <- function() {
  suppressMessages(
    suppressWarnings(
      vpc::read_table_nm(med_20mb_file, perl = T)
      )) 
}
med_rnm <- function() {
  suppressMessages(
    suppressWarnings(
      read_nonmem(med_20mb_file)
      )) 
}
med_xpose <- function() {
  suppressMessages(
    suppressWarnings(
    xpose4::read.nm.tables(med_20mb_file)
    ))
}

```

```{r}
large_rtnm <- function() {
  suppressMessages(
    suppressWarnings(
      vpc::read_table_nm(large_120mb_file, perl = T)
      )) 
}
large_rnm <- function() {
  suppressMessages(
    suppressWarnings(
      read_nonmem(large_120mb_file)
      )) 
}
large_xpose <- function() {
  suppressMessages(
    suppressWarnings(
    xpose4::read.nm.tables(large_120mb_file)
    ))
}
```

```{r}
bm_med <- microbenchmark(
  med_rtnm(),
  med_rnm(),
  med_xpose(),
  times = 5L
)
```

```{r}
bm_large <- microbenchmark(
  large_rtnm(),
  large_rnm(),
  large_xpose(),
  times = 5L
)
```

```{r}
bm_med
bm_large
```



